The 2014 SMT Competition

### 2015-08-03 2014/report/SMTCOMP.tex
Undefined behavior explained.

\paragraph{Undefined Behavior.} During the testing period prior to the competition beginning, different solvers exhibited different behavior on some benchmarks. Investigation revealed that these differences were the result of different treatments of partial operations, divide by zero, in particular. The question is what should be the definition of divide by zero in SMT-LIB's logic of total functions. This question had been discussed at length over previous years, with proposals to clarify the SMT-LIB standard.
There was certainly not time to resolve this matter prior to the competition, not to mention correcting solvers to the agreed behavior.\footnote{The discussion this time around can be found in the archives of the \url{smtcomp-discussion@lists.sourceforge.net} mailing list, in the few days near 2014-06-16.}
Thus, for the competition, we decided to omit all benchmarks (6060 of them, 1348 of which were otherwise eligible) that were affected by divide-by-zero behavior. This question, however, remains open.

@@

\paragraph{Competition Timeline}

@@

The competition used the NSF-funded StarExec~\cite{DBLP:conf/cade/StumpST14} computational cluster at U.~Iowa for executing solvers on benchmarks. Past instances of SMT-COMP used a previous SMT-Exec infrastructure. StarExec was used for several other competitions in 2014 as well.

StarExec consists of 192 2.4\,GHz computational nodes, running Red Hat Linux~6.3. Each node has a dual-processor, quad-core CPU (model Intel Xeon CPU E5-2609 at 2.4GHz), 10240 KB cache, and either 129022\,MB (32 nodes) or 258294\,MB of memory (160 nodes). In competition, processes were limited to 100\,GB memory. To avoid interference among job-pairs, only one job-pair was executed at a time on each node. The software configuration contains
\begin{itemize}[noitemsep,nolistsep]
\item Linux kernel 2.6.32-573.1.1.el6.x86\_64
\item GNU C Library stable release version 2.12
\item gcc (GCC) 4.4.7 20120313 (Red Hat 4.4.7-11)
\end{itemize}

@@

Based on the experience of 2014, the 2014 organizers have the following recommendations or topics for consideration for future competitions.
\begin{itemize}
\item The 2014 competition used all available benchmarks; though the computation resources enable using all benchmarks, future competitions should consider how to make a principled selection to avoid over-representing benchmarks of particular types or origin.
\item The competition has used numbers of solved benchmarks as the primary success criterion. Time to solve benchmarks should be considered more strongly. In particular, a separate track that emphasizes fast solution of fairly simple problems might be informative.
\item Related to the previous point, currently if a solver times-out or issues a response of `unknown', the time taken to do so is not counted in the accumulated time. Only the time taken to compute correct responses is counted towards the evaluation metric.
A user's experience, however, is that the time taken for a solver to say ``I don't know'' is just as important as time to produce a useful answer. Thus we recommend that such
computation time be included in the evaluation metric. Omitting this time has had no effect so far because time was little used in the overall metric.
\item The comparison to a previous years' results will now be easier because common hardware will be used and benchmark selection is simplified. Future organizers might also include a larger selection of the specific solver versions that were entered in previous competitions. 
\item A key improvement needed is better benchmark sets. Though the accumulation of benchmark problems since the inception of the competition is impressive, attention now needs to be paid to the quality and distribution of benchmarks. Some divisions are represented by only a few benchmarks; others have large numbers of similar benchmarks.
Benchmarks representative of application scenarios are particularly important.
\item Subsequent to SMT-COMP 2014, David Cok, Aaron Stump, Morgan Deters, and Clark Barrett collaborated in resolving the status of many previously unknown benchmarks. Those new expected results are not yet included in the SMT-LIB benchmarks on StarExec. They should be incorporated into SMT-LIB on StarExec prior to the next competition.
\item If a global metric is needed again in the future (per \S\ref{sec:floc}), a review and re-discussion of the appropriate metric should be instigated.
\item Solvers with breadth of application across many logics and solvers that address problems in new logics such as string and floating point computations are important to users. Future competitions should add tracks or otherwise find means to reward solvers that implement such capabilities.
\item Various competition tracks used in the past should be rejuvenated: parallel processing, computation of unsat cores, production of proofs, computation of symbolic models, and computation of concrete counterexamples.
\item One missing tool is a standard SMT-LIB syntax checker. jSMTLIB~\cite{cok2011jsmtlib} has been proposed for this purpose, but is not yet integrated into StarExec. The tool also needs to be updated to include the proposed new SMT-LIB features.
\item The difficulty in preparing a new solver for submission to StarExec for participation in SMT-COMP should be reduced.
\item Resolve the definition of divide-by-zero behavior in SMT-LIB.
\item An item for study is the effect of benchmark scrambling on the outcome of solver comparisons.
\item Clearly define the subset of SMT-LIB~v2 that solvers must support for a competition and that benchmarks must use.
\end{itemize}

>>>>

\paragraph{Undefined behavior.} \label{par:undefined} During the testing period prior to the competition beginning, different solvers exhibited different behavior on some benchmarks. Investigation revealed that these differences were the result of different treatments of partial operations; in particular, divide by zero.  The semantics of SMT-LIB is based on a logic of total functions, meaning that division by zero cannot raise an exception (or the like), but must have \emph{some} value.  The SMT-LIB standard leaves this value unspecified (i.e., it could be anything).  In contrast, some solvers assume a specific, fixed value.  This question had been discussed at length over previous years, with proposals to clarify the SMT-LIB standard.  There was certainly not time to resolve this matter prior to the competition, not to mention correcting solvers to the agreed behavior.\footnote{The discussion this time around can be found in the archives of the \url{smtcomp-discussion@lists.sourceforge.net} mailing list, in the few days near 2014-06-16.}  Thus, for the competition, we decided to omit all benchmarks (6060 of them, 1348 of which were otherwise eligible) that were potentially affected by divide-by-zero behavior.  This question, however, remains open.

@@

\paragraph{Competition timeline.}

@@

The competition used the NSF-funded StarExec~\cite{DBLP:conf/cade/StumpST14} computational cluster at the University of Iowa for executing solvers on benchmarks. Past instances of SMT-COMP used a previous SMT-Exec infrastructure. StarExec was used for several other competitions in 2014 as well.

StarExec currently consists of 192 2.4\,GHz computational nodes, running Red Hat Linux~6.3. Each node has a dual-processor, quad-core CPU (model Intel Xeon CPU E5-2609), 10240~KB cache, and either 129022\,MB (32 nodes) or 258294\,MB of memory (160 nodes). In competition, processes were limited to 100\,GB memory. To avoid interference among job-pairs, only one job-pair was executed at a time on each node. The software configuration contains
\begin{itemize}[noitemsep,nolistsep]
\item Linux kernel 2.6.32-573.1.1.el6.x86\_64
\item GNU C Library stable release version 2.12
\item gcc (GCC) 4.4.7 20120313 (Red Hat 4.4.7-11)
\end{itemize}

@@

Based on the experience of 2014, the 2014 organizers have the following recommendations or topics for consideration for future competitions.
\begin{itemize}
\item The 2014 competition used all available benchmarks; though the computation resources enable using all benchmarks, future competitions should consider how to make a principled selection to avoid over-representing benchmarks of particular types or origin.
\item The competition has used numbers of solved benchmarks as the primary success criterion. Time to solve benchmarks should be considered more strongly. In particular, a separate track that emphasizes fast solution of fairly simple problems might be informative.
\item Related to the previous point, currently if a solver times-out or issues a response of `unknown', the time taken to do so is not counted in the accumulated time. Only the time taken to compute correct responses is counted towards the evaluation metric.
A user's experience, however, is that the time taken for a solver to say ``I don't know'' is just as important as time to produce a useful answer. Thus we recommend that such
computation time be included in the evaluation metric. Omitting this time has had no effect so far because time was little used in the overall metric.
\item The comparison to a previous years' results will now be easier because common hardware will be used and benchmark selection is simplified. Future organizers might also include a larger selection of the specific solver versions that were entered in previous competitions. 
\item A key improvement needed is better benchmark sets. Though the accumulation of benchmark problems since the inception of the competition is impressive, attention now needs to be paid to the quality and distribution of benchmarks. Some divisions are represented by only a few benchmarks; others have large numbers of similar benchmarks.
Benchmarks representative of application scenarios are particularly important.
\item Subsequent to SMT-COMP 2014, David Cok, Aaron Stump, Morgan Deters, and Clark Barrett collaborated in resolving the status of many previously unknown benchmarks. Those new expected results are not yet included in the SMT-LIB benchmarks on StarExec. They should be incorporated into SMT-LIB on StarExec prior to the next competition.
\item If a global metric is needed again in the future (per \S\ref{sec:floc}), a review and re-discussion of the appropriate metric should be instigated.
\item Solvers with breadth of application across many logics and solvers that address problems in new logics such as string and floating point computations are important to users. Future competitions should add tracks or otherwise find means to reward solvers that implement such capabilities.
\item Various competition tracks used in the past should be rejuvenated: parallel processing, computation of unsat cores, production of proofs, computation of symbolic models, and computation of concrete counterexamples.
\item One missing tool is a standard SMT-LIB syntax checker. jSMTLIB~\cite{cok2011jsmtlib} has been proposed for this purpose, but is not yet integrated into StarExec. The tool also needs to be updated to include the proposed new SMT-LIB features.
\item The difficulty in preparing a new solver for submission to StarExec for participation in SMT-COMP should be reduced.
\item Resolve the definition of divide-by-zero behavior in SMT-LIB (cf.~\S\ref{par:undefined}).
\item An item for study is the effect of benchmark scrambling on the outcome of solver comparisons.
\item Clearly define the subset of SMT-LIB~v2 that solvers must support for a competition and that benchmarks must use.
\end{itemize}
